---
title: "Project_1: How old the mussel is?"
author: 
date: "10/26/2020"
output: html_document
html_document:
    toc: true
    toc_depth: 3log(27, base = 3) * (sqrt(144) + 10)
    toc_float: true
    number_section: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(cowplot)
theme_set(theme_bw())
```

# Prepearing Data

Сombining all observations into a one table

```{r merging files}

merging_data = function(mypath){
  filenames=list.files(path=mypath, full.names=TRUE)
  datalist = lapply(filenames, function(x){read.csv(file=x,header=T)})
  Reduce(function(x,y) {merge(x,y,all = TRUE)}, datalist)
}

mussels_all = merging_data(mypath)
```

Let's look at the data structure:

```{r data structure}

str(mussels_all)

```

You can see that not all variables are of the correct type: Rings should be numeric rather than character.
Also, using the "unique" function, we can see that there are NA and other incorrectly entered values here.

Fixing the Rings variables values:

```{r Rings variables structure}

mussels_all$Rings <- as.numeric(mussels_all$Rings)
unique(mussels_all$Rings)

```

Fixing the Sex variables values. Let's change the name that is too long, and there are 3 levels here, NA, and misnamed parameters, we can find out the positions of the incorrectly named lines and rename them:

```{r Sex variables structure}

names(mussels_all)[2] <- "Sex"
unique(mussels_all$Sex)

match('male',mussels_all$Sex)
mussels_all[1122, "Sex"] <- "1"

match('one',mussels_all$Sex)
mussels_all[4176, "Sex"] <- "1"

match('three',mussels_all$Sex)
mussels_all[3488, "Sex"] <- "3"

unique(mussels_all$Sex)

mussels_all$Sex <- factor(mussels_all$Sex, levels = c(1, 2, 3), labels = c("Male", "Female", "Uvenil"))

```

The Length variables:

```{r Length variables structure}

unique(mussels_all$Length)
match('No data! I forgot to mesure it!(',mussels_all$Length)
mussels_all[3487, "Length"] <- "NA"
mussels_all$Length <- as.numeric(mussels_all$Length)

```

Now, let's see the structure of dataframe:

```{r Dataframe structure}

summary(mussels_all)

```

We can see, that there are some NA values, but there are not very many of them, so I think we can just delete them

```{r NA removal}

mussels <- na.omit(mussels_all)

```

Check the data again:

```{r data check}

summary(mussels)
str(mussels)
colSums(is.na(mussels))

```

# EDA analysis

Now everything seems to be fine and we can go to EDA analysis:

Firstly, we are going to visualize the variables for outliers

```{r outliers check, echo = FALSE}

library(ggplot2)
library(cowplot)
Rings_box <- ggplot(mussels, aes(x = factor(Sex), y = Rings))+
     geom_boxplot()
Length_box <- ggplot(mussels, aes(x = factor(Sex), y = Length))+
  geom_boxplot()
Diameter_box <- ggplot(mussels, aes(x = factor(Sex), y = Diameter))+
  geom_boxplot()
Height_box <- ggplot(mussels, aes(x = factor(Sex), y = Height))+
  geom_boxplot()
Whole_weight_box <- ggplot(mussels, aes(x = factor(Sex), y = Whole_weight))+
  geom_boxplot()
Shucked_weight_box <- ggplot(mussels, aes(x = factor(Sex), y = Shucked_weight))+
  geom_boxplot()
Viscera_weight_box <- ggplot(mussels, aes(x = factor(Sex), y = Viscera_weight))+
  geom_boxplot()
Shell_weight_box <- ggplot(mussels, aes(x = factor(Sex), y = Shell_weight))+
  geom_boxplot()
plot_grid(Rings_box, Length_box, Diameter_box, Height_box, Whole_weight_box, Shucked_weight_box, Viscera_weight_box, Shell_weight_box)

```

Next we should assess the presence of relationships between variables:

```{r correlation check}
pairs(mussels)
mussels_numeric <- mussels[c(1,3:9)]
cor(mussels_numeric)

```

## What hypotheses can be formulated?

Ho: There is no relationship between age, mussel size and number of rings.
Ha: There is a relationship between the number of rings and the size and age of the mussel.

## Calculation the mean and standard deviation of the Length variable for molluscs of different sexes:

```{r mean and SD}

library(dplyr)
MeanSD_Length <- cbind(aggregate(Length ~ Sex, mussels, mean), aggregate(Length ~ Sex, mussels, sd)) 
MeanSD_Length

```

## Calculation for what percentage of molluscs the value of the Height variable does not exceed 0.165?

```{r Smallest mussels}

mussels_less_then165 <- mussels %>% filter(Height < 0.165)
mussels_less_then165_in_percent <- length(mussels_less_then165$Height) / length(mussels$Height) * 100
mussels_less_then165_in_percent

```

## Calculate what the Length variable is equal to, which is greater than 92% of all observations

```{r Lengthest mussels}

quantile(mussels$Length, 0.92)

```

## Length values after standardization:

```{r Z scale}
library(dplyr)
mussels <- mussels %>% mutate(Lenght_z_scores = (Length - mean(Length)) / sd(Length))
```

## Comparison of the diameter of mussels with the number of rings 5 and 15

```{r five and fifteen rings}

library(dplyr)
comparisson_btw_5_and_15_rings <- rbind(mussels %>% filter(Rings == 5), mussels %>% filter(Rings == 15))

```

## Check the data for normality. If the data distribution is normal, we can use the t-test to compare the diameter of mussels with 5 and 15 rings

```{r five and fifteen rings2}

shapiro.test(comparisson_btw_5_and_15_rings$Length[comparisson_btw_5_and_15_rings$Rings == 5])
shapiro.test(comparisson_btw_5_and_15_rings$Length[comparisson_btw_5_and_15_rings$Rings == 15])


diameter_ttest <- t.test(Diameter ~ Rings, comparisson_btw_5_and_15_rings)

```

Comparison results in the graph form:

```{r five and fifteen rings3, echo = FALSE}

ggplot(comparisson_btw_5_and_15_rings, aes(x = factor(Rings), y = Diameter))+
  geom_boxplot() +
  xlab(label = "Number of rings") + 
  ylab(label = "Diameter")

```

Thus, the alternative hypothesis is correct, with an increase in the number of rings, the diameter of the mussel increases.

## Estimating Diameter and Whole_weight variables

Check the data for normality:

```{r Diameter and Whole_Weight}

shapiro.test(mussels$Diameter)
shapiro.test(mussels$Whole_weight)

cor.test(x = mussels$Diameter, y = mussels$Whole_weight, method = 'spearman')

```

The distribution data is not entirely linear, however, we see the relationship between the diameter and weight of the mussel

```{r Diameter and Whole_Weight2, echo = FALSE}

ggplot(mussels, aes(x = Diameter, y = Whole_weight, col = Sex))+
  geom_point(alpha = 0.7, size = 2)

```

## Conclusion
Thus, we see that the older the mussel, the more rings it has and the more pain it is in size and weight.















